### 2020-09-14 paper/appendA.tex
Much pogrress

\begin{defn}[Artemis graph]
  A graph $G$ is an \emph{Artemis graph} if it contains no odd hole, no antihole of length at least five, and no prism.
\end{defn}

@@

\begin{defn}[subdivision]
  In a graph $G$, subdividing an edge $uv \in E(G)$ means removing the edge $uv$ and adding a new vertex $w$ and two new edges $uw, wv$. Starting with a graph $G$, the effect of repeatedly subdividing edges prioduces a graph $H$ called a \emph{subdivision of $G$}. Note that $V(G) \subseteq V(H)$.
\end{defn}

\begin{defn}[triad]
  In a graph $G$, a \emph{triad} is a set of three pairwise non-adjacent vertices.
\end{defn}

\begin{defn}[good partition]
  A \emph{good partition} of a graph $G$ is a partition $(K_1, K_2, K_3, L, R)$ of $V(G)$ such that:
  \begin{itemize}
    \item $L$ and $R$ are not empty, and $L$ is anticomplete to $R$,
    \item $K_1 \cup K_2$ and $K_2 \cup K_3$ are cliques,
    \item in the graph obtained from $G$ by removing all edges between $K_1$ and $K_3$, every chordless path with one end in $K_1$, the other in $K_3$, and interior in $L$ contains a vertex from $L$ that is complete to $K_1$,
    \item either $K_1$ is anticomplete to $K_3$, or no vertex in $L$ has neighbors in both $K_1$ and $K_3$,
    \item for some $x \in L$ and $y \in R$, there is a triad of $G$ that contains $\{x, y\}$.
  \end{itemize}
\end{defn}

@@

\graphAppendix{Color Graph}
\input{coloringAppendix/algI.tex}
\clearpage

\graphAppendix{Color Good Partition}
\input{coloringAppendix/algII.tex}
\clearpage

\graphAppendix{Grow Hyperprism}
\input{coloringAppendix/algIII.tex}
\clearpage

\graphAppendix{Good Partition from an even hyperprism}
\input{coloringAppendix/algIV.tex}
\clearpage

\graphAppendix{Good Partition from odd hyperprism}
\input{coloringAppendix/algV.tex}
\clearpage

\graphAppendix{Good Partition from $J$-strip system}
\input{coloringAppendix/algVIa.tex}
\clearpage

\graphAppendix{Good Partition from a special $K_4$ strip system}
\input{coloringAppendix/algVIb.tex}
\clearpage

>>>>

\begin{defnTwo}[Artemis graph]
  A graph $G$ is an \emph{Artemis graph} if it contains no odd hole, no antihole of length at least five, and no prism.
\end{defnTwo}

@@

\begin{defnTwo}[subdivision]
  In a graph $G$, subdividing an edge $uv \in E(G)$ means removing the edge $uv$ and adding a new vertex $w$ and two new edges $uw, wv$. Starting with a graph $G$, the effect of repeatedly subdividing edges prioduces a graph $H$ called a \emph{subdivision of $G$}. Note that $V(G) \subseteq V(H)$.
\end{defnTwo}

\begin{defnTwo}[triad]
  In a graph $G$, a \emph{triad} is a set of three pairwise non-adjacent vertices.
\end{defnTwo}

\begin{defnTwo}[good partition]
  A \emph{good partition} of a graph $G$ is a partition $(K_1, K_2, K_3, L, R)$ of $V(G)$ such that:
  \begin{itemize}
    \item $L$ and $R$ are not empty, and $L$ is anticomplete to $R$,
    \item $K_1 \cup K_2$ and $K_2 \cup K_3$ are cliques,
    \item in the graph obtained from $G$ by removing all edges between $K_1$ and $K_3$, every chordless path with one end in $K_1$, the other in $K_3$, and interior in $L$ contains a vertex from $L$ that is complete to $K_1$,
    \item either $K_1$ is anticomplete to $K_3$, or no vertex in $L$ has neighbors in both $K_1$ and $K_3$,
    \item for some $x \in L$ and $y \in R$, there is a triad of $G$ that contains $\{x, y\}$.
  \end{itemize}
\end{defnTwo}

@@

\graphAppendix{Color square-free perfect graph}
\input{coloringAppendix/algI.tex}
\clearpage

\graphAppendix{Color good partition}
\input{coloringAppendix/algII.tex}
\clearpage

\graphAppendix{Grow hyperprism}
\input{coloringAppendix/algIII.tex}
\clearpage

\graphAppendix{Good partition from an even hyperprism}
\input{coloringAppendix/algIV.tex}
\clearpage

\graphAppendix{Good partition from an odd hyperprism}
\input{coloringAppendix/algV.tex}
\clearpage

\graphAppendix{Good partition from a $J$-strip system}
\input{coloringAppendix/algVIa.tex}
\clearpage

\graphAppendix{Good partition from a special $K_4$ strip system}
\input{coloringAppendix/algVIb.tex}
\clearpage

### 2020-09-14 paper/coloringBergeGraphs.tex
Much pogrress

Consider an "umbrella" in $\mathbb{R}^3$ with the unit vector $e_1 = (1, 0, 0)$ as its "handle" and 5 "ribs" of unit length \todo{picture}. Open it up to the point where non-consecutive ribs are orthogonal, that is form an angle of 90$^\circ$. This way we get a representation of $C_5$ by 5 unit vectors $u_1, \ldots u_5$ so that each $u_i$ forms the same angle with $e_1$ and any two non-adjacent nodes are represented with orthogonal vectors. We can calculate $e_1^\intercal u_i = 5 ^ {-1/4}$.

@@

\begin{alg}[maximum cardinality stable set in a perfect graph]
  \label{alg:maxStableSet}
  Input: A perfect graph $G = (V, E)$.

@@

\begin{alg}[maximum weighted stable set in a perfect graph]
  \label{alg:maxWeightedStableSet}
  Input: A perfect graph $G = (V, E)$ and a weight function $w : V \rightarrow \mathbb{N}$.

@@

Recall that to calculate a maximum cardinality clique of $G$, we can calculate a maximum cardinality stable set of $\overline{G}$.
\begin{alg}
  \label{alg:ssIntersectingCliques}
  Input: A perfect graph $G = (V, E)$.

@@

\begin{alg}
  \label{alg:minColoring}
  Input: A perfect graph $G = (V, E)$.

@@

A most recent advancement (2018) is an algorithm by Chudnovsky et al. that colors any square-free Berge graphs in time of $O(|V|^9)$ \cite{Chudnovsky2019}. Before proving strong perfect graph conjecture, a similar conjecture for square-free Berge graphs has been proven by Conforti et al. \cite{Conforti2004} During one of her lectures, Maria Chudnovsky expressed hope that discovery of full algorithm for coloring Berge graphs might follow a similar pattern. We analyze this algorithm and provide its pseudocode in \cref{ch:coloringSquareFree}.

\TODO{Dałbym ze 2-3 zdania braggowania, ze algorytm jest duzo bardziej zlozony niz rozpoznawanie + na czym sie opiera.}

>>>>

\begin{wrapfigure}{l}{0.4\textwidth}
  \input{tikzpictures/umbrella.tex}
  \caption{Lovász's umbrella}
  \label{fig:umbrella}
\end{wrapfigure}

Consider an "umbrella" in $\mathbb{R}^3$ with the unit vector $e_1 = (1, 0, 0)$ as its "handle" and 5 "ribs" of unit length.  Open it up to the point where non-consecutive ribs are orthogonal, that is form an angle of 90$^\circ$. This way we get a representation of $C_5$ by 5 unit vectors $u_1, \ldots u_5$ so that each $u_i$ forms the same angle with $e_1$ and any two non-adjacent nodes are represented with orthogonal vectors. See \Cref{fig:umbrella}. We can calculate $e_1^\intercal u_i = 5 ^ {-1/4}$.

@@

\begin{alg}[Maximum cardinality stable set in a perfect graph]
  \label{alg:maxStableSet}
  Input: A perfect graph $G = (V, E)$.

@@

\begin{alg}[Maximum weighted stable set in a perfect graph]
  \label{alg:maxWeightedStableSet}
  Input: A perfect graph $G = (V, E)$ and a weight function $w : V \rightarrow \mathbb{N}$.

@@

Recall that to calculate a maximum cardinality clique of $G$, we can calculate a maximum cardinality stable set of $\overline{G}$.
\begin{alg}[Stable set intersecting all maximum cardinality cliques]
  \label{alg:ssIntersectingCliques}
  Input: A perfect graph $G = (V, E)$.

@@

\begin{alg}[Color perfect graph]
  \label{alg:minColoring}
  Input: A perfect graph $G = (V, E)$.

@@

A most recent advancement (2018) is an algorithm by Chudnovsky et al. that colors any square-free Berge graphs in time of $O(|V|^9)$ \cite{Chudnovsky2019}. Before proving strong perfect graph conjecture, a similar conjecture for square-free Berge graphs has been proven by Conforti et al. \cite{Conforti2004}. During one of her lectures, Maria Chudnovsky expressed hope that discovery of full algorithm for coloring Berge graphs might follow a similar pattern.

The algorithm for coloring square-free Berge graphs is much more complex than CCLSV algorithm for Berge graphs recognition and would be much harder to implement. The main idea is that, if square-free Berge graph $G$ is a member of class that is simpler to analyze and color (Artemis graphs), then it must have an induced prism. We find it and grow it until it is maximal and then we can create a partition of $G$, that we color recursively. The paper in which the algorithm is presented \cite{Chudnovsky2019} is of much more mathematical ''flavor'' then \cite{MC05} -- there is a proof that a $O(|V|^9)$ algorithm exists, but actually deducing it from the paper is far from straightforward. We do it in the Appendix A.

### 2020-09-14 paper/experiments.tex
Much pogrress

\begin{figure}
  \centering
  \input{plots/detailed.csv.detailed.pgf}
  \caption{Detailed running times}
  \label{plot:splitLines}
\end{figure}

When looking at the results (\Cref{plot:perfLines}), while na\"ive algorithm's time is almost zero, we see the CCLSV running time climbing with the growth of $|V|$, with only moderate improvement by GPU CCLSV. Slight GPU improvement is explained by anylysis of the components of the overall time (\Cref{plot:perfDet}). Testing all near cleaners takes up around half of total time, so we cannot speed up the overall time by much. These are also relatively small tests, so the latency from copying data to the GPU is significant.

Next, we generated a random bipartite graphs (see paragraph above) and then calculated their line graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficieng number of graphs for each $|V|$. On \cref{plot:perf2Lines} we can see na\"ive algorithm being much faster on the majority of the tests, although its faster growth allows GPU CCLSV to be faster on the biggest tests.

We turn our attention to graphs generated on a grid, or a checker board (see later paragraphs). First, we take lattice graphs. In lattice graph, each vertex is connected to vertices that are above, below, to the left and to the right of it, if such vertices exist. See \cref{fig:lattice} for an example of a $4 \times 4$ lattice graph.

Let us take a look at \cref{plot:gridLines}. It is a clear example of why polynomial algorithms can be useful, even the ones with time complexity as high as CCLSV. For $|V| = 48$ the running time for the CCLSV is 22.2s, for GPU CCLSV is 7.7s and for na\"ive is 1.5s. Na\"ive is almost 15x faster. But with growing $|V|$ its running time grows very fast, reaching 41s for $|V| = 60$ and 210s for $|V| = 66$. CCLSV growth is much gentler and for $|V| = 66$ it is 2.5x faster than na\"ive.

If we imagine our grid to be a chessboard we can define a rook graph to have an edge, when the rook from chess can move from one vertex to another in one move. See \cref{fig:rook} for an example of $4 \times 4$ Rook graph.

Rook graphs have many more chords than Lattice graphs, so the na\"ive's performance is much better here. What is more interesting, the overall performance of CCLSV is much worse (\cref{plot:rookLines}). Looking at \cref{plot:rookDet} we can see that the majority of CCLSV time is spent checking all near cleaners. This can be explained by the fact that rook graphs are much denser than lattice graphs and have many more 3-vertex paths (see line TODO of \cref{alg:testNearCleaner}). This is confirmed by greater CCLSV GPU speedup of around 6x on graphs with $|V| = 35$.

Hypercube graphs tell a similar story to lattice graphs and knight graphs. Na\"ive algorithm is very fast at first, but becomes unbearably slow for $|V|$ higher than 55. Unfortunately, CCLSV doesn't fare much better, although it grows slower. Again, the speedup of GPU CCLSV is noticeable, being around 4.8x for $|V| = 55$.

Split graphs we use are unions of cliques and independent sets of same size, with some edges between them. In generating them, each edge between a vertex from a clique and a vertex from an independent set has had $1/2$ chance to appear.

There are no long paths here, so na\"ive algorithm works great (\cref{plot:splitLines}). What is interesting, when looking at detailed times of GPU CCLSV, we can see that checking the existence of simple forbidden structures takes more than 30\% of the time. This opens a possibility of further GPU speedups, not explored by us.

>>>>

\begin{figure}
  \centering
  \input{plots/detailed.csv.detailed.pgf}
  \caption{Detailed running times}
  \label{plot:perfDetailed}
\end{figure}

First, let us take a look at the performance of CCLSV algorithm. If we limit runtime to about two minutes, we can run it on graphs of sizes up to 40-70, depending on the type of the graph we test. What is interesting, that there isn't that much of a variance of the running time depending on the type of graph we test on. This is a big advantage of CCLSV over the na\"ive algorithm, which is highly dependent on the type of input. On some graphs, it works almost instantenously, what is to be expected, given its nature of enumerating paths. But on some graphs, e.g. lattice graphs (\Cref{plot:gridLines}) its overall running time grows much faster than CCLSV's.

Let us turn our attention to the GPU CCLSV's results. \Cref{plot:perfDetailed} shows breakdown of each implementation running times on a single graph of each class. We can clearly see that in most cases, testing possible near-cleaners takes most of the running time, so the speedup by the GPU is significant.

Examples of full binary graph, and lattice graph show that getting near cleaners still can be a big bottleneck. Parallelization of this method could be considered in the future. A possible idea would be to use recently developed dynamic dictionary for the GPU \cite{Ashkiani2018}. When running GPU CCLSV on split graphs, quite a lot of time is spent checking simple structures. This should be easily parallelizable if need be, or if graphs that require it are encountered.

% When looking at the results (\Cref{plot:perfLines}), while na\"ive algorithm's time is almost zero, we see the CCLSV running time climbing with the growth of $|V|$, with only moderate improvement by GPU CCLSV. Slight GPU improvement is explained by anylysis of the components of the overall time (\Cref{plot:perfDet}). Testing all near cleaners takes up around half of total time, so we cannot speed up the overall time by much. These are also relatively small tests, so the latency from copying data to the GPU is significant.

% Next, we generated a random bipartite graphs (see paragraph above) and then calculated their line graphs. This gives us graphs of varying $|V|$, so we repeated the process until we had sufficieng number of graphs for each $|V|$. On \cref{plot:perf2Lines} we can see na\"ive algorithm being much faster on the majority of the tests, although its faster growth allows GPU CCLSV to be faster on the biggest tests.

% We turn our attention to graphs generated on a grid, or a checker board (see later paragraphs). First, we take lattice graphs. In lattice graph, each vertex is connected to vertices that are above, below, to the left and to the right of it, if such vertices exist. See \cref{fig:lattice} for an example of a $4 \times 4$ lattice graph.

% Let us take a look at \cref{plot:gridLines}. It is a clear example of why polynomial algorithms can be useful, even the ones with time complexity as high as CCLSV. For $|V| = 48$ the running time for the CCLSV is 22.2s, for GPU CCLSV is 7.7s and for na\"ive is 1.5s. Na\"ive is almost 15x faster. But with growing $|V|$ its running time grows very fast, reaching 41s for $|V| = 60$ and 210s for $|V| = 66$. CCLSV growth is much gentler and for $|V| = 66$ it is 2.5x faster than na\"ive.

% If we imagine our grid to be a chessboard we can define a rook graph to have an edge, when the rook from chess can move from one vertex to another in one move. See \cref{fig:rook} for an example of $4 \times 4$ Rook graph.

% Rook graphs have many more chords than Lattice graphs, so the na\"ive's performance is much better here. What is more interesting, the overall performance of CCLSV is much worse (\cref{plot:rookLines}). Looking at \cref{plot:rookDet} we can see that the majority of CCLSV time is spent checking all near cleaners. This can be explained by the fact that rook graphs are much denser than lattice graphs and have many more 3-vertex paths (see line TODO of \cref{alg:testNearCleaner}). This is confirmed by greater CCLSV GPU speedup of around 6x on graphs with $|V| = 35$.

% A hypercube is a generalization of a cube to a higher dimensional space. We use hypercubes that are not complete, to achieve higher granularity of data. In a hypercube of $n$ vertices there is an edge between vertices $u$ and $v$ if and only if binary representations of $u$ and $v$ differ by exactly one bit.

% Hypercube graphs tell a similar story to lattice graphs and knight graphs. Na\"ive algorithm is very fast at first, but becomes unbearably slow for $|V|$ higher than 55. Unfortunately, CCLSV doesn't fare much better, although it grows slower. Again, the speedup of GPU CCLSV is noticeable, being around 4.8x for $|V| = 55$.

% Split graphs we use are unions of cliques and independent sets of same size, with some edges between them. In generating them, each edge between a vertex from a clique and a vertex from an independent set has had $1/2$ chance to appear.

% There are no long paths here, so na\"ive algorithm works great (\cref{plot:splitLines}). What is interesting, when looking at detailed times of GPU CCLSV, we can see that checking the existence of simple forbidden structures takes more than 30\% of the time. This opens a possibility of further GPU speedups, not explored by us.

### 2020-09-14 paper/implementation.tex
Much pogrress

Another crucial optimization is to eliminate partial solutions that have chords right away. This is especially useful in graphs that have many chords and reduces running time dramatically.

@@

This optimization alone is crucial for performance. In our testing, all calls of $areUnique$ took abour 70\% of total running time of the CCLSV algorithm, and virtually all running time of the na\"ive algorithm \todo{check}. After using static $stamp$ array, it fell to 5.2\% in CCLSV and to X\% in na\"ive. The overall speedup of the CCLSV algorithm was about 6x. \todo{check if for sure it didn't use path speedup} The speedup of na\"ive algorithm was even greater -- about 20x.

@@

Because of a good performance of our optimized algorithm for generating paths, we use it whenever possible. It can be easily modified to generate all possibly chordal paths or holes and therefore has much use in the CCLSV algorithm. For example, when searching for jewels we generate with it all possibly chordal paths of length 5 and for each check if it has required properties of a jewel. This proved to be much faster than generating vertices that don't have forbidden chords one by one. We use the same algorithm for generating starting vertices for a possible $\T_2$ and $\T_3$.

@@

\TODO{example of a call graph?}

\TODO{callgrind}
\TODO{first result to best result gains (like: we are 20x faster than what first comes to mind)}
\TODO{regenerate speedup results to make sure what the gains were}

@@

\TODO{CUDA background}
\TODO{moderngpu - allows us to run simple transforms}
\TODO{Implementing get NC is tough -- we rely on set<bitset> to do the work. Maybe we could use a dynamic set from Adam's link? As a "open question"}

We posed a question if it would be profitable to use GPU for this problem. On one hand the graphs we are working on are small and our time complexity so large that a speedup from massively parallel architecture could be profitable. On the other the algorithm is complex and not easy to parallelise. Therefore we decided to analyze the  CCLSV\todo{a good name for it?} algorithm and parrallelize parts that would benefit from it the most.

@@

It turns out that this optimization alone speeds up \cref{alg:testNearCleaner} almost 7x \todo{check} on bigger tests, which gives us a speedup of about 6x \todo{check} for overall algorithm.

\TODO{implement other CUDA opts and describe them}

>>>>

Another crucial optimization is to eliminate partial solutions that have edges between not consecutive vertices right away. Also, when we look for an odd hole specifically, we can assume that its first vertex we enumerate is the lowest. Those two optimizations alone make a huge difference. With them, our na\"ive algorithm implementation went from being able to check graphs with $|V|$ up to 12-14, within a couple of minutes, to being faster than CCLSV on most of our tests.

@@

This optimization almost cuts down na\"ive algorithm's running time in half.

@@

Because of a good performance of our optimized algorithm for generating paths, we use it whenever possible. It can be easily modified to generate all possibly chordal paths or holes and therefore has much use in the CCLSV algorithm. For example, when searching for jewels we generate with it all possibly chordal paths of length 5 and for each check if it has required properties of a jewel. This proved to be much faster than generating vertices that don't have forbidden chords one by one. We use the same algorithm for generating starting vertices for a possible $\T_2$ and $\T_3$. Before optimizing path enumeration algorithm, all calls of $areUnique$ alone took abour 70\% of total running time of the CCLSV algorithm. After optimizing $areUnique$, path enumeration still took more than 50\% of total running time. After all optimizations it is almost unnoticable.

@@

@@

CUDA is a parrallel computing platform model created by Nvidia and executed on Nvidia GPUs. It allows to run a program in parallel by many CUDA threads at once. Although usually described by PRAM models, implementing PRAM algorithm in CUDA architecture is not as straightforward as implementing RAM algorith on a cpu. Some of the causes are: synchronization between threads is very costly, there are multiple memory models to choose from, threads run divided in blocks and it is possible to efficiently share memory within blocks and copying memory from and back to CPU suffers from large latency. What is more, there are no standard dynamic structures in PRAM model, let alone CUDA, that are taken for granted in PRAM model, such as hashing maps or sets. We used a library called \emph{moderngpu} for our implementation, because it allows a slightly cleaner code.

\TODO{Implementing get NC is tough -- we rely on set<bitset> to do the work. Maybe we could use a dynamic set from Adam's link? As a "open question"}

We posed a question if it would be profitable to use GPU for the problem of recognizing Berge graphs. On one hand the graphs we are working on are small and our time complexity so large that a speedup from massively parallel architecture could be profitable. On the other the algorithm is complex and not easy to parallelise. Therefore we decided to analyze the  CCLSV algorithm and parrallelize parts that would benefit from it the most.

@@

The part of the CCLSV implemented on the gpu achieved speedup of up to 15x. Which gives us a speedup of up to 9x for overall algorithm.

### 2020-09-14 paper/perfectGraphs.tex
Much pogrress

\begin{lemma}[Roussel-Rubio Lemma \cite{RR01,MC05}]\label{lem:Roussel-Rubio}
	Let $G$ be Berge, $X$ be an anticonnected subset of $V(G)$, $P$ be an odd path $p_1\ldots p_n$ in $G\setminus X$ with length at least 3, such that $p_1$ and $p_n$ are $X$-complete and $p_2, \ldots, p_{n-1}$ are not. Then:
	\begin{itemize}
		\item $P$ is of length at least 5 and there exist nonadjacent $x, y \in X$, such that there are exactly two edges between $x, y$ and $P^*$, namely $xp_2$ and $yp_{n-1}$,
		\item or $P$ is of length 3 and there is an odd antipath joining internal vertices of $P$ with interior in $X$.
	\end{itemize}
	\TODO{We may use this lemma quite often, might want to provide proof if so.}
\end{lemma}

@@

\begin{alg}
	\label{alg:t3}
	Input: A graph $G$.

@@

	Each execution of lines \ref{line:t3n4start}-\ref{line:t3mid} takes $O(|V|^2)$ time, and there are $O(|V|)$ $v_4$s to enumerate. Line \ref{line:t3v4if} takes $O(|V|)$ time and then there are $O(|V|)$ $v_3$s to enumerate. Each is checked in $O(1)$ time. If we find the correct one, we calculate a path $P$, but return right away, so we calculate $P$ at most once in whole algorithm.
\end{proof}

@@

\begin{defn}[C-major vertices]
  Given a shortest odd hole $C$ in $G$, a node $v \in V(G) \setminus V(C)$ is $C$-major if the set of its neighbors in $C$ is not contained in any 3-node path of $C$.
  \TODO{a picture of this, clean odd hole, amenable hole}
\end{defn}

@@

\begin{defn}[amenable odd hole]
  An odd hole $C$ of $G$ is \emph{amenable} if it is a shortest off hole in $G$, it is of length at least 7 and for every anticonnected set $X$ of $C$-major vertices there is a $X$-complete edge in $C$.
\end{defn}

@@

The proof of this theorem is quite technical and we will not discuss it here. See section 8 of \cite{MC05} for the proof.

>>>>

\begin{lemma}[Roussel-Rubio Lemma \cite{RR01,MC05}]\label{lem:Roussel-Rubio}
	Let $G$ be Berge, $X$ be an anticonnected subset of $V(G)$, $P$ be an odd path $p_1\ldots p_n$ in $G\setminus X$ with length at least 3, such that $p_1$ and $p_n$ are $X$-complete and $p_2, \ldots, p_{n-1}$ are not. Then:
	\begin{itemize}
		\item $P$ is of length at least 5 and there exist nonadjacent $x, y \in X$, such that there are exactly two edges between $x, y$ and $P^*$, namely $xp_2$ and $yp_{n-1}$,
		\item or $P$ is of length 3 and there is an odd antipath joining internal vertices of $P$ with interior in $X$.
	\end{itemize}
\end{lemma}

@@

\begin{alg}[Test if $G$ contains a $\T_3$]
	\label{alg:t3}
	Input: A graph $G$.

@@

	Each execution of lines \ref{line:t3n4start}-\ref{line:t3mid} takes $O(|V|^2)$ time, and there are $O(|V|)$ $v_4$s to enumerate. Line \ref{line:t3v4if} takes $O(|V|)$ time and then there are $O(|V|)$ $v_3$ to enumerate. Each is checked in $O(1)$ time. If we find the correct one, we calculate a path $P$, but return right away, so we calculate $P$ at most once in whole algorithm.
\end{proof}

@@

\begin{defn}[C-major vertices]
  Given a shortest odd hole $C$ in $G$, a node $v \in V(G) \setminus V(C)$ is $C$-major if the set of its neighbors in $C$ is not contained in any 3-node path of $C$.
\end{defn}

@@

\begin{defn}[amenable odd hole]
  An odd hole $C$ of $G$ is \emph{amenable} if it is a shortest odd hole in $G$, it is of length at least 7 and for every anticonnected set $X$ of $C$-major vertices there is a $X$-complete edge in $C$.
\end{defn}

\begin{figure}
  \begin{minipage}{.5\textwidth}
      \input{tikzpictures/amenable.tex}
			\captionof{figure}{Amenable odd hole}
			\label{fig:amenableOddHole}
  \end{minipage}%
  \begin{minipage}{.5\textwidth}
      \input{tikzpictures/notAmenable.tex}
			\caption{Not amenable odd hole}
			\label{fig:notAmenableHole}
  \end{minipage}
  \end{figure}

Some pictures might be helpful. \Cref{fig:amenableOddHole} shows an example of an amenable odd hole $C = v_1, \ldots, v_7$. It is not a clean hole -- vertices $v_8$ and $v_9$ are $C$-major, and $v_8, v_9, v_1, v_2, v_3$ is an axample of a near-cleaner for $C$.
\Cref{fig:notAmenableHole} shows an odd hole $C = v_1, \ldots, v_7$ that is not amenable, because there is no $\{v_8, v_9\}$-complete edge in $C$.

@@

The proof of this theorem is quite technical and we will not discuss it here. See section 8 of \cite{MC05} for the proof. We can see an example of the contraposition of the \Cref{thm:amenableHoles} working in \Cref{fig:notAmenableHole} -- vertices $v_1$, $v_2$, $v_8$, $v_6$, $v_7$ and a path $v_1-v_9-v_5-v_6$ form a jewel.
